---
title: "Gator Exploration"
output: html_document
date: "2022-10-24"
---

```{r setup, include=FALSE, message=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(quantreg)
library(lqmm)
library(googlesheets4)
```
Ryan Shipley smoothed out the data and calculated corrected CO2 and O2 fits, as well as quantiles for the metabolic rates that are included in the 'data_summary.csv' file that we'll work with. This has the various quantiles calculated for each 5 minute observation period, so there are multiple observations per animal. If we want to calculate SMR in a simlar way to other studies, then we will need to look for the lowest, stable metabolic rate for 10, 15, or 30 minutes (whatever has been used before).  

If we want to do any additional exploration of parameters like RER, then we'll go back into Ryan's code and re-calculate that for our individuals. 

Let's look at some data.
```{r summary data}
gatorsOG <- read.csv("./Modified_Data/data_summary.csv")
gatorsOG$id <- as.character(gatorsOG$id)
gatorExpInf <- read.csv("./Raw_Data/FMS Experiment Data Collection.xlsx - Sheet1.csv")
massID <- gatorExpInf %>% select(id = AnalysisAnimalID, mass = Mass, svl = SVL)
gators <- gatorsOG %>% left_join(massID)
names(gators)
```

Plots. What we can see below is that 2 animals (91 & 113) have *much* higher metabolic rates. But also that the quantiles within an individual are pretty similar to one another, so even though there was some noise in the data it's not too crazy.

```{r}
gators %>% pivot_longer(cols = starts_with("X"),
                       names_to = "quantile",
                       values_to = "VO2") %>% #Turn the data longwise to make facets easier
   arrange(treatment, id) %>% #hopefully organize by treatment
   ggplot()+
   geom_boxplot(aes(x = as.character(id), group = as.character(id), y = VO2, fill = treatment))+
  theme_bw()+
  scale_fill_discrete(name="Experimental\nTreatment",
                         breaks=c("C", "MT"),
                         labels=c("Control", "MT"))+
  labs(x = "animal ID", y = "VO2 - ml O2 / min")+
   facet_wrap(~quantile, ncol = 2)
```

We can calculate the variance across the quantiles to see this. It is at or near 0 for all of the five minute intervals.

```{r}
gators %>% pivot_longer(cols = starts_with("X"),
                       names_to = "quantile",
                       values_to = "VO2") %>% 
  group_by(id, treatment, min5_interval) %>% 
  mutate(quantVar = var(VO2)) %>% 
  ggplot()+
  geom_density(aes(x = quantVar, fill = treatment), alpha = 0.5)+
  theme_bw()+
  scale_fill_discrete(name="Experimental\nTreatment",
                         breaks=c("C", "MT"),
                         labels=c("Control", "MT"))+
  labs(x = "among-quantile variance")
```


####Size variation

Just a look at mass & SVL. There is a pretty good distribution between the treatment groups in size, but 2 animals stick out. 91 is a chonker and also had really high MR in the plots above. 113 is the next heaviest and also had a high metabolic rate. This & in combination with their very high resting metabolic rates give me some reason to exclude them from subsequent analysis. This is going to drop the sample size of the control animals to 6. 

```{r}
ggplot(gators)+
  geom_point(aes(x = svl, y = mass, color = treatment))+
  geom_text(aes(x = svl, y = mass, label = id), data =gators, 
            nudge_x = 0.05, nudge_y = 0.25, check_overlap = F)+
  scale_color_discrete(name="Experimental\nTreatment",
                         breaks=c("C", "MT"),
                         labels=c("Control", "MT"))+
  labs(x = "SVL (cm)", y = "mass (g)")+
  theme_bw()
```


###Analysis of MR & Treatment, Body Mass
Filter out 91 & 113 for analysis. Then run a glmm on median values

```{r}
dropgator <- c("113", "91")
ok_gator <- gators %>% filter(!(id %in% dropgator))
library(lme4)
library(car)
library(MuMIn)

m1 <-lmer(X0.5~treatment*mass+(1|id), data = ok_gator) 
summary(m1)

```
The summary of the model shows that the random effect of id has relatively large variance. When we look at the Fixed Effects table, we can see that the estimates for treatment & mass are both negative, but they have a positive interaction. This doesn't tell us the 'significance' of those terms but does show us the magnitude of the effects. We can use 'Anova' from the *car* package to look at the significance of these terms using a likelihood ratio model.


```{r}
Anova(m1)
```

There are no significant effects of either mass or treatment on alligator median metabolic rates.

So overall, looking at the median VO2 values there isn't an effect of treatment on metabolic rates. The only effect we see is that of mass -- and larger alligators have higher metabolic rates. 

We can make a plot of mass vs VO2 that accounts for differences among individuals

```{r}
ok_gator %>% 
ggplot()+
  geom_boxplot(aes(x = mass, y = X0.5, group = id, fill = treatment))+
  theme_bw()+
  scale_fill_discrete(name="Experimental\nTreatment",
                         breaks=c("C", "MT"),
                         labels=c("Control", "MT"))+
  labs(x = "mass (g)", y = "median VO2 - ml O2 / min")

```

### Standard Metabolic Rate
Previous work has used the minimum MR in a 15 minute period. We can use our lowest quantile and see if there is any effect there. There isn't, which isn't much of a surprise considering how little variance there was among the quantiles.

```{r}
m2 <-lmer(X0.05~treatment*mass+(1|id), data = ok_gator) 
summary(m2)
Anova(m2)
```
Plot the lowest quantile / standard metabolic rate

```{r}
ok_gator %>% 
ggplot()+
  geom_boxplot(aes(x = mass, y = X0.05, group = id, fill = treatment))+
  theme_bw()+
  scale_fill_discrete(name="Experimental\nTreatment",
                         breaks=c("C", "MT"),
                         labels=c("Control", "MT"))+
  labs(x = "mass (g)", y = "SMR: VO2 - ml O2 / min")
```





